---
title: "Customizing plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(ggxpose)

xpdb <- xpdb_ex_pk

knitr::opts_chunk$set(fig.dpi = 96,
                      fig.align = 'center', 
                      fig.height = 4, 
                      fig.width = 4,
                      out.width = '75%')
```


### Plot type
The option `type` allows to control how the data should be represented ie. whether to use points `'p'`, lines `'l'`, smooth `'s'` or any combination of the three (eg. `type = 'pl'`).

```{r demo type, echo = FALSE, fig.width = 6, out.width = '100%'}
gridExtra::grid.arrange(
  dv_vs_ipred(xpdb, title = "type = \'p\'", subtitle = FALSE, caption = FALSE, type = 'p'),
  dv_vs_ipred(xpdb, title = "type = \'l\'", subtitle = FALSE, caption = FALSE, type = 'l'),
  dv_vs_ipred(xpdb, title = "type = \'s\'", subtitle = FALSE, caption = FALSE, type = 's'),
  ncol = 3)
```


### Visual guides
The option `guides` can enable (`TRUE`) or disable (`FALSE`) the visual guides on the graphs such as unity line on the `dv_vs_ipred()` plot or horizontal lines on `cwres_vs_time()`.


### Titles
All ggxpose plots have by default an informative title, subtitle and caption. For example all plots using individual model predictions (IPRED) will display the shrinkage of the epsilons. These titles can easily be edited as templates using keywords wich will be replaced by their actual value stored in the summary level of the `xpdb` object when rendering the plots. Keywords are defined by a word preceeded by a `@` e.g. `'@ofv'`. A list of all available keyword can be accessed via `help('template_titles')`. Finally, the title, subtitle or caption can be disabled by setting them to `FALSE`.

```{r demo titles}
# Using template titles
dv_vs_ipred(xpdb,
            title    = 'DV vs. IPRED (@run, obj: @ofv)',
            subtitle = 'Based on: @nind subjects and @nobs records',
            caption  = 'Dir: @dir')

# Disabling all titles
dv_vs_ipred(xpdb, title = FALSE, subtitle = FALSE, caption = FALSE)
```


### Modify aesthetics
By default the aesthetics are read from the `xp_theme` level in the xpdb object but these can be modified in any plot function. ggxpose makes use of the `ggplot2` functions mapping for any layer (e.g. points, lines, etc.) however to direct the mapping to a specific layer, a prefix appealing to the targeted layer should be used. The format is defined as `layer_aesthetic = value`. Hence to change the color of points in ggplot2 the argument `color = 'green'` could be used in `geom_point()`, while in ggxpose the same could be achieved with `point_color = 'green'`.

In basic goodness-of-fit plots, the layers have been named as: `point_xxx`, `line_xxx`, `smooth_xxx`, `guide_xxx`, `panel_xxx`, `xscale_xxx`, `yscale_xxx` where `xxx` can be any option available in the `ggplot2` layers: `geom_point`, `geom_line`, `geom_smooth`, `geom_abline`, `facet_wrap`, `scale_x_continuous`, etc.

```{r demo aesthetics}
dv_vs_ipred(xpdb, 
            # Change points aesthetics
            point_color = 'blue', point_alpha = 0.5, 
            point_stroke = 0, point_size = 1.5, 
            # Change lines aesthetics 
            line_alpha = 0.5, line_size = 0.5, 
            line_color = 'orange', line_linetype = 'solid', 
            # Change smooth aesthetics
            smooth_method = 'lm')
```

Aesthetics can also be defined mapped to a variable using the ggplot2 `aes()` function.
```{r demo mapping}
dv_vs_ipred(xpdb, type = 'p',
            ggplot2::aes(point_color = as.factor(SEX))) 
```


### Grouping variable
The argument group defines the grouping variable to be used in the dataset __for the lines only__, by default group is defined to the column `'ID'`. To apply a grouping variable on any other layer, a manual mapping can be made using the argument `xxx_group = variable`


### Panels
Panels (or faceting) can be created by using the `by` argument as follows.

```{r demo panels}
dv_vs_ipred(xpdb, by = 'SEX')
```

Additional arguments can be passed to the panel function using the `panel_xxx` notation (eg. `panel_ncol = 2`, `panel_labeller = 'label_both'`). For more options read the help of [`facet_wrap`](http://docs.ggplot2.org/current/facet_wrap.html) from `ggplot2`.


### Additional layers
`ggxpose` offers the opportunity to add any additional layers from `ggplot2`. Hence, a `ggplot2::geom_rug()` layer could be added to the `dv_vs_ipred()` plot along with some annotations (`ggplot2::annotate()`). Note: the additional layers do not inherit from the ggxpose aesthetic mapping (i.e. colors or other options need to be defined in each layer as shown below).

Layers can also be used to modify the aesthetics scales for example  `ggplot2::scale_color_manual()`, or remove a legend `ggplot2::scale_fill_identity()`.

```{r demo layers}
dv_vs_ipred(xpdb) +
  geom_rug(alpha = 0.2, color = 'grey50',
           sides = 'lb', size = 0.4) +
  annotate(geom = 'text', 
           color = 'darkred',
           label = 'LLOQ', x = -4, y = -4) +
  annotate(geom = 'rect',
           alpha = 0.2, fill = 'red',
           xmin = -Inf, xmax = -3,
           ymin = -Inf, ymax = -3)
```

### Scales options
The argument `log` allows to automatically log the relevant axes when set to `TRUE`. Additional arguments can be provided to the scales via the mapping by using the naming convention `xscale_xxx` or `yscale_xxx` where `xxx` is the name of a `ggplot2` scale argument such as `name`, `breaks`, `labels`, `expand`.

```{r scales demo}
dv_vs_ipred(xpdb, 
            xscale_breaks = c(-4, -2, 0),
            xscale_labels = c('Low', 'Med', 'High'),
            xscale_expand = c(0.2, 0),
            xscale_name = 'Individual model prediction')
```


### Themes
Theme in `ggxpose` are easily customizable. Themes are made up of two parts:

 - A `ggplot2` (`gg_theme`) theme controlling the aspect of the plot window    
 - An `xpose` (`xp_theme`) theme controlling the aspect of the layers such as points or lines

Themes can be attached to an `xpdb` when importing the data with `xpose_data()`, using the function `xpose_theme()` on an xpdb object or when creating a graph.

```{r demo themes xpdb, eval = FALSE}
# While creating the xpdb
xpdb <- xpose_data(runno = '001', 
                   gg_theme = theme_classic(), 
                   xp_theme = xp_theme_xpose4())

# Update a pre-existing xpdb
xpdb <- xpose_theme(xpdb     = xpdb,
                    gg_theme = theme_bw(),
                    xp_theme = c(point_color = 'dodgerblue4',
                                 line_color  = 'dodgerblue4')) 
```
                                 
```{r demo themes plot}
# For a single plot only
dv_vs_ipred(xpdb        = xpdb, 
            gg_theme    = ggplot2::theme_dark(), 
            point_color = 'darkblue')
```


### Examples of `gg_theme` themes:

*Note: `theme_readable()` was created to make the plot more readable by increasing labels size. The visibility of the grids was also reduced to bring the focus on the data itself.*


### Examples of `xp_theme` themes:

*Note: `gg_theme = theme_bw()` was also used in the xpose4 plot*


### The pipes
`ggxpose` makes use of the pipe operator `%>%`. It can be used to generate clear flow as follows.

```{r, eval = FALSE}
xpose_data(runno = '001') %>% 
  dv_vs_ipred() %>% 
  xpose_save()
```
