---
title: "Introduction"
author: "Benjamin Guiastrennec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
library(xpose)

xpdb <- xpdb_ex_pk %>% 
set_var_types(idv = 'TIME')

knitr::opts_chunk$set(fig.dpi = 96,
fig.align = 'center', 
fig.height = 4, 
fig.width = 4,
out.width = '50%',
comment = '',
message = FALSE)
```


### Load xpose
The first step is to load xpose in R with the following command:
```r
library(xpose)
```

### Import model output
The function `xpose_data()` collects all model output files and table and organizes them into an R object commonly called `xpdb` which stands for "xpose database".

``` r
xpdb <- xpose_data(runno = '001', dir = 'analysis/model/pk/')
```

### Glimpse at the xpdb
The files attached to an xpdb object can be displayed to the console simply by writing its name to the console or by using the `print()` function.

```{r demo print xpose_data}
xpdb # or print(xpdb)
```

### Model summary
A summary of a model run can be displayed to the console by using the `summary()` function on an xpdb object.

```{r demo xpose summary}
summary(xpdb)
```


### Parameter estimates
A table of parameter estimates can be displayed to the console by using the `prm_table()` function on an xpdb object.

```{r demo prm_table}
prm_table(xpdb)
```


### Listing variables
A list of available variables for plotting can be displayed to the console by using the `list_vars()` function on an xpdb object.

```{r demo list_vars}
list_vars(xpdb)
```

### Pipes
`xpose` makes use of the pipe operator `%>%` from the package [dplyr](http://dplyr.tidyverse.org). Pipes can be used to generate clear workflow.

```{r, eval = FALSE}
xpose_data(runno = '001') %>% 
  dv_vs_ipred() %>% 
  xpose_save()
```

### Editing the xpdb
Multiples edits can be made to the xpdb object. For instance the type (visible using the `list_vars()` function described above) of a variable can be changed. Hence the independent variable (idv) could be changed from `TIME` (default in NONMEM) to `TAD`. All plots using `idv` will then automatically use `TAD`.

```{r, change idv}
# With the TIME default
xpdb %>% 
  dv_vs_idv()

# After IDV reassignment
xpdb %>% 
  set_var_types(idv = 'TAD') %>% 
  dv_vs_idv()
```

### Generating plots
Plotting functions are used as follows:
``` {r demo basic gof}
# DV vs. IPRED plot
dv_vs_ipred(xpdb)

# CWRES vs. PRED plot
res_vs_pred(xpdb, res = 'CWRES')
```


### Saving plots
The `xpose_save` function was designed automatically save plots according to their run number (`@run`) and the plotting function (`@plotfun`) using keywords preceded by a `@`. You can learn more about these keywords using `help('template_titles')`. The file extension is guessed from the file name and must match one of .pdf, .jpeg, .png, .bmp or .tiff, if missing a pdf will be generated by default. Finally, if the `plot` argument is left empty `xpose_save` will automatically save the last plot that was created or modified.

```{r, eval = FALSE}
# Save the last plot
xpose_save()

# Change file name and extension
xpose_save(file = '@run_@plotfun_[@ofv].jpeg', dir = 'pk/diagnostics')
```
